### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6922fbcfc88b17b9b42b7d43e59e81efebaa73cb 🚀

Bats, the second most diverse groups of mammals, are found worldwide and
serve as the main animal reservoir for different strains of
betacoronaviruses.\textsuperscript{13} This has attracted attention to
areas where high diversity of bats, and therefore presumably high
diversity of betacoronaviruses, can be an important issue for human
health.\textsuperscript{14,15} By overlaying the IUCN range maps for
confirmed bat hosts of betacoronaviruses {[}fig.~\ref{fig:richness};
top{]}, we see that the the main hotspots (both in terms of size and
higher values) of host richness are primarily South-Eastern Asia, parts
of Southern Europe, and to a lesser extent parts of Africa in the -25-0
range of latitudes. The description of host richness is an important
first step towards understanding risk, as previous
research\textsuperscript{2,16} states that locally diverse bat
communities could maintain more viruses and hence, a higher probability
of having a pathogen that could represent a risk for human health.

@@

As host richness, joint distinctiveness, or phylogeographic structure
suggest that the bat-betacoronaviruses complex is globally fragmented
enough to give rise to both different levels of risk (as evidenced by
the spatial location of spillover events) and different types of
co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
Coevolution\textsuperscript{7} to provide a measure of risk accounting
for multiple processes. In fig.~\ref{fig:trivariate}, we overlapped
three components of spillover risk: viral sharing, \emph{i.e.} the
chance that two bats will share viruses overall; Local Contribution to
Beta Diversity, \emph{i.e.} the fact that a bat community is
compositionally unique compared to the average compositional similarity
across the entire system; finally, host phylogenetic diversity,
\emph{i.e.} how dispersed the bats in a location are within the tree of
life. This approach leads to the definition of broad biogeographic
regions of risk, where the same color represents the same type of risk.
By way of constrat to figures fig.~\ref{fig:richness} and
fig.~\ref{fig:biogeo}, these regions do not necessarilly overlap with
previous spatial partitions of the bat-betacoronaviruses complex.

@@

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronaviruses
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In
fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
areas, which are essentially limited to South-Eastern Asia, and to some
part of Sub-Saharan Africa. Adopting a geographic mosaic theory
perspective on risk, other regions of the world are of lesser concern
fig.~\ref{fig:risk}. Our risk decomposition does not account for viral
diversity or distinctiveness. The simple rationale behind it is that the
acquisition of viral data is rarely disconnected from the acquisition of
host data. There are more sources of information on hosts than on
viruses, allowing to develop a host-centric perspective on risk
(although this estimate would more accurate with viral traits related to
\emph{e.g.} ability to switch hosts or pathogenic potential). Areas with
high bat diversity and high turnover \emph{may} facilitate the
evolutionary radiation of viruses, matching previous findings that the
diversification of bat coronaviruses is driven largely by host shifts
(inter-genus or higher levels of cross-species transmission) and, to a
lesser degree, cospeciation and sharing.\textsuperscript{intra-genus
cross-species transmission; 2} This diversification is not an actual
risk factor for spillover itself, but acts downstream of a spillover
event by increasing the random chance of the emergence of a virus with
the raw genomic components required for the potential to infect humans.

@@

From another perspective, areas of high host uniqueness and virus
sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
risk through mixing of unique viruses (via codivergence) and in turn
recombination. Under our framework, such a hotspot was identified in
Madagascar, where most bat species are endemic following evolutionary
divergence from sister species in both African and Asian
continents.\textsuperscript{\emph{e.g.} 36} Recent
surveillance\textsuperscript{37} has identified a novel
\emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus},
\emph{Rousettus madagascariensis}), emphasizing strong proof of
principle in model predictions.

@@

Based on the previous result, we extracted the risk component from the
composite map (see Methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}. As
this map represents the potential risk, it must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more
widely.\textsuperscript{38} Since human activity is required to amplify
the frequency of virus encounters and thus create areas of viral
amplification, mapping the potential risk against measures of land use
is required to generate a more actionable assessment of risk. This map
is presented in fig.~\ref{fig:compound}. Most of South America and
Europe are at comparatively lower risk, as although densely populated,
settlements tend to be in areas with lower potential risk. Regions like
Malaysia and the North coast of Australia have a high risk component,
but should represent a relatively lower effective risk due to low human
density. However, this mapping reveals that South-East Asia, the Indian
subcontinent, and parts of sub-Saharan Africa, are at high risk due to
the overlap between built areas and bat communities representing more
opportunities for cross-species transmission of betacoronaviruses. In
looking for the origins of SARS in China,\textsuperscript{39} present
serological evidence that strongest human-animal contact results in
higher risk of virus exposure, regardless of the animal species, but
that different types of contact had different impacts. Ideally,
finer-grained information about human activity (rather than human
presence through anthropisation) could allow to partition this risk
further, alebit at the cost of more hypotheses required to estimate the
amount of risk represented by each activity. Our map of risk overlays
with recent results from\textsuperscript{40} -- areas of purported high
risk/diversitifcation potential (Madagascar, South-America) overlay with
sampling gaps for \emph{Betacoronavirus}, stressing the need for
spatially targeted monitoring and discovery.

>>>>

Bats, the second most diverse groups of mammals, are found worldwide and
serve as the main animal reservoir for different strains of
betacoronaviruses.\textsuperscript{13} This has attracted attention to
areas where high diversity of bats, and therefore presumably high
diversity of betacoronaviruses, can be an important issue for human
health.\textsuperscript{14,15} By overlaying the IUCN range maps for
confirmed bat hosts of betacoronaviruses (fig.~\ref{fig:richness}; top),
we see that the the main hotspots (both in terms of size and higher
values) of host richness are primarily South-Eastern Asia, parts of
Southern Europe, and to a lesser extent parts of Africa in the -25-0
range of latitudes. The description of host richness is an important
first step towards understanding risk, as previous
research\textsuperscript{2,16} states that locally diverse bat
communities could maintain more viruses and hence, a higher probability
of having a pathogen that could represent a risk for human health.

@@

As host richness, joint distinctiveness, or phylogeographic structure
suggest that the bat-betacoronaviruses complex is globally fragmented
enough to give rise to both different levels of risk (as evidenced by
the spatial location of spillover events) and different types of
co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
Coevolution to provide a measure of risk accounting for multiple
processes. In fig.~\ref{fig:trivariate}, we overlapped three components
of spillover risk: viral sharing, \emph{i.e.} the chance that two bats
will share viruses overall; Local Contribution to Beta Diversity,
\emph{i.e.} the fact that a bat community is compositionally unique
compared to the average compositional similarity across the entire
system; finally, host phylogenetic diversity, \emph{i.e.} how dispersed
the bats in a location are within the tree of life. This approach leads
to the definition of broad biogeographic regions of risk, where the same
color represents the same type of risk. By way of constrat to figures
fig.~\ref{fig:richness} and fig.~\ref{fig:biogeo}, these regions do not
necessarilly overlap with previous spatial partitions of the
bat-betacoronaviruses complex.

@@

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronaviruses
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In
fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
areas, which are essentially limited to South-Eastern Asia, and to some
part of Sub-Saharan Africa. Adopting a geographic mosaic theory
perspective on risk, other regions of the world are of lesser concern
(fig.~\ref{fig:risk}). Our risk decomposition does not account for viral
diversity or distinctiveness. The simple rationale behind it is that the
acquisition of viral data is rarely disconnected from the acquisition of
host data. There are more sources of information on hosts than on
viruses, allowing to develop a host-centric perspective on risk
(although this estimate would more accurate with viral traits related to
\emph{e.g.} ability to switch hosts or pathogenic potential). Areas with
high bat diversity and high turnover \emph{may} facilitate the
evolutionary radiation of viruses, matching previous findings that the
diversification of bat coronaviruses is driven largely by host shifts
(inter-genus or higher levels of cross-species transmission) and, to a
lesser degree, cospeciation and sharing, representing intra-genus
cross-species transmission.\textsuperscript{2} This diversification is
not an actual risk factor for spillover itself, but acts downstream of a
spillover event by increasing the random chance of the emergence of a
virus with the raw genomic components required for the potential to
infect humans.

@@

From another perspective, areas of high host uniqueness and virus
sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
risk through mixing of unique viruses (via codivergence) and in turn
recombination. Under our framework, such a hotspot was identified in
Madagascar, where most bat species are endemic following evolutionary
divergence from sister species in both African and Asian
continents.\textsuperscript{36} Recent surveillance\textsuperscript{37}
has identified a novel \emph{Betacoronavirus} (in the subgenus
\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
strong proof of principle in model predictions.

@@

Based on the previous result, we extracted the risk component from the
composite map (see Methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}. As
this map represents the potential risk, it must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more
widely.\textsuperscript{38} Since human activity is required to amplify
the frequency of virus encounters and thus create areas of viral
amplification, mapping the potential risk against measures of land use
is required to generate a more actionable assessment of risk. This map
is presented in fig.~\ref{fig:compound}. Most of South America and
Europe are at comparatively lower risk, as although densely populated,
settlements tend to be in areas with lower potential risk. Regions like
Malaysia and the North coast of Australia have a high risk component,
but should represent a relatively lower effective risk due to low human
density. However, this mapping reveals that South-East Asia, the Indian
subcontinent, and parts of sub-Saharan Africa, are at high risk due to
the overlap between built areas and bat communities representing more
opportunities for cross-species transmission of betacoronaviruses. In
looking for the origins of SARS in China,\textsuperscript{39} present
serological evidence that strongest human-animal contact results in
higher risk of virus exposure, regardless of the animal species, but
that different types of contact had different impacts. Ideally,
finer-grained information about human activity (rather than human
presence through anthropisation) could allow to partition this risk
further, alebit at the cost of more hypotheses required to estimate the
amount of risk represented by each activity. Our map of purported high
risk/diversitifcation potential (Madagascar, South-America) overlay with
sampling gaps for \emph{Betacoronavirus},\textsuperscript{40} stressing
the need for spatially targeted monitoring and discovery.
